---
title: "Data Analytics Assignment"
subtitle: "Pre-primary school doesn't matter?"
author: "Angélina Gardin"
format: html
editor: visual
execute: 
  echo: false
  warning: false
  message: false
theme: Zephyr
backgroundcolor: "floralwhite"
---
<center>

# Pre-primary school doesn't matter?

</center>

<br> 

[**95,87%**]{style="color: darkred;"} This is the out-of-school rate for children one year younger than official entry age that **Yemen** is currently facing.

[**0.14%**]{style="color: darkred;"} This is the out-of-school rate for children one year younger than official entry age that **Ireland** is currently facing.

<br> 

------------------------------------------------------------------------

<br> 

**Is that a coincidence?** Why there is a such significant difference between this two numbers?  Is pre-primary school only **optional** and mainly related to cultural or contextual **habits**? What is the real impact of that on children?

Maybe are you thinking that waiting for official entry age in primary school is *sufficient*. Let me show you the real consequences of this phenomenon. Please have a look on the map below.

<br> 

```{r}
library(tidyverse)
options(scipen = 999)
library(readxl)
data <- read_csv("C:/Users/angel/OneDrive/Documents/Neoma/DCU/Data Analytics and Story Telling/Assignment/unicef_metadata.csv", 
                 col_types = cols(year = col_number(), 
                                  `Population, total` = col_number(), 
                                  `GDP per capita (constant 2015 US$)` = col_number(), 
                                  `GNI (current US$)` = col_number()))
data<-select(data,country,year,pop="Population, total",gdp="GDP per capita (constant 2015 US$)",gni="GNI (current US$)")
ind <- read_csv("C:/Users/angel/OneDrive/Documents/Neoma/DCU/Data Analytics and Story Telling/Assignment/unicef_indicator_1.csv", 
                col_types = cols(time_period = col_number(), 
                                 obs_value = col_number()))
ind<-select(ind,country,year=time_period,value=obs_value,sex)

literacy <- read_excel("C:/Users/angel/OneDrive/Documents/Neoma/DCU/Data Analytics and Story Telling/Assignment/GLOBAL_DATAFLOW_1970-2024.xlsx", 
                       col_types = c("text", "text", "text", 
                                     "numeric", "numeric", "numeric", 
                                     "text", "numeric", "numeric", "numeric", 
                                     "numeric", "numeric", "numeric", 
                                     "numeric", "text", "numeric", "numeric", 
                                     "numeric", "numeric", "numeric", 
                                     "text"))
literacy<-select(literacy,country="Geographic area",sex=Sex,year=TIME_PERIOD,lrate=OBS_VALUE)
completion <- read_excel("C:/Users/angel/OneDrive/Documents/Neoma/DCU/Data Analytics and Story Telling/Assignment/GLOBAL_DATAFLOW_1970-2024 (1).xlsx", 
                         col_types = c("text", "text", "text", 
                                       "numeric", "numeric", "numeric", 
                                       "text", "numeric", "numeric", "numeric", 
                                       "numeric", "numeric", "numeric", 
                                       "numeric", "text", "numeric", "numeric", 
                                       "numeric", "numeric", "numeric", 
                                       "text"))
completion <- select(completion,country="Geographic area",sex=Sex,year=TIME_PERIOD,crate=OBS_VALUE)
join<-full_join(data,ind,by=c("country"="country","year"="year"))

join<-full_join(join,literacy,c("country"="country","sex"="sex","year"="year"))
join<-full_join(join,completion,c("country"="country","sex"="sex","year"="year"))
map<- map_data("world")
map$region<-ifelse(map$region=="USA","United States",map$region)
join<-full_join(join,map,c("country"="region"))
joinout<-filter(join,sex=="Total",value!="NA")
ggplot(joinout)+
  aes(long,lat,group=group,fill=value)+
  geom_polygon()+
  theme_classic()+
  scale_fill_gradient(low = "lightgoldenrod1", high = "tomato3", na.value = "grey")+
  labs(
    title = "Current out-of-school rate for children one year youger than official entry age",
    subtitle = "(Countries with no value are not displayed on the map)",
    caption = "Source: Unicef",
    x="Longitude",
    y="Latitude",
    fill = "Out-of-school rate (%)")
```

------------------------------------------------------------------------

**What did you see?** Have you noticed that countries with the lower rate seems to be countries known as emerging countries? So, what do you think right now? The out-of-school rate seems related to the economic situation of the country. You are not sure yet? Let's check if there is an effective correlation between the economic situation of a country and its out-of-school rate for children on year younger than official entry age.

<br> 

```{r}
joingdp<-filter(join,value!="NA",sex=="Total")
ggplot(joingdp)+
  aes(x = gdp,y=value,color=country)+
  geom_point()+
  guides(color ="none")+
  labs(x="Current GDP per capita (US$)", y="Out-of-school rate (%)", title="Evolution of the out-of-school rate regarding the GDP per capita", caption="Unicef")+
  theme_classic()+
  geom_smooth(method = "lm", color="black")+
  scale_x_continuous(breaks = c(25000, 50000, 75000,100000))+
  scale_y_continuous(breaks = c(0,25,50,75,100))
```

<br> 

**No doubt remaining:** the out-of-school rate for children one year younger than official entry age significantly depend on the economic situation of the country. Poor countries have poor access or use of pre-primary education. But what are the long-term consequences for children?

<br>

**First consequence:** the lack of pre-primary education directly impact the completion rate for adolescents of lower secondary school age as you can see with the *negative correlation* between the two indicators in the graphs below.

```{r}
outvscompl<-right_join(ind,data,by=c("country"="country","year"="year"))
outvscompl<-select(outvscompl,-year)
outvscompl<-full_join(outvscompl,completion,by=c("country"="country","sex"="sex"))
outvscompl<-filter(outvscompl,sex=="Total",value!="NA",crate!="NA")
outvscompl<-filter(outvscompl,country%in%c("Thailand","Rwanda","Argentina","Costa Rica","Uruguay","Guatemala","Angola","Chad"))

ggplot(outvscompl)+
  aes(reorder(country,value),value, fill="darkseagreen")+
  geom_col()+
  labs(x=NULL,
       y="Out-of-school rate (%)",
       title="Out-of-school rate per country",
       caption="Unicef",
       theme_classic())+
  guides(fill ="none")



ggplot(outvscompl)+
  aes(reorder(country,value),crate, fill="darkseagreen")+
  geom_col()+
    labs(x=NULL,
         y="Completion rate (%)",
         caption="Unicef",
         theme_classic(),
       title="Completion rate per country")+
  guides(fill ="none")

```

<br>

**Second consequence:** all the educational system is affected by the lack of pre-primary education. Countries with higher out-of school rate for primary education tend to be countries where the literacy of young adults is also the lowest. This is what we can observe in those bar chart comparing the out-of-school rate to the literacy rate for the same countries.

```{r}
outvslit<-right_join(ind,data,by=c("country"="country","year"="year"))
outvslit<-select(outvslit,-year)
outvslit<-full_join(outvslit,literacy,by=c("country"="country","sex"="sex"))
outvslit<-filter(outvslit,sex=="Total",value!="NA",lrate!="NA")
outvslit<-filter(outvslit,country%in%c("Viet Nam","Niger","Guinea","Benin","Pakistan","Togo"))
ggplot(outvslit)+
  aes(reorder(country,value),value,fill="lightsalmon")+
  geom_col()+
    labs(x=NULL,
         y="Out-of-school rate (%)",
         caption="Unicef",
         theme_bw(),
       title="Out-of-school rate per country")+
  guides(fill ="none")
ggplot(outvslit)+
  aes(reorder(country,value),lrate,fill="lightsalmon")+
  geom_col()+
    labs(x=NULL,
         y="Literacy rate (%)",
         caption="Unicef",
         theme_classic(),
       title="Literacy rate per country")+
  guides(fill ="none")

```

***
<br>

**Definitely**, consequences of the out-of-school rate are rude for countries. But let's bring us hope for the future. Indeed, with the constant evolution of global GDP over the years, we can expect a development of the economy in most countries over the world and by consequent expect a fall in the out-of-school rate for the coming decades.

```{r}
evolgdp<-select(data,year,country,gdp)
evolgdp<-filter(evolgdp,-gdp!="NA",year>=2002&year<=2019)
evolgdp<-evolgdp %>%
  group_by(year) %>%
  summarise(gdp=sum(gdp))
ggplot(evolgdp)+
  aes(x=year,y= gdp,color="royalblue")+
  geom_line(arrow = arrow())+
  labs(caption="Unicef",
       x= NULL,
       y="Global GDP (US$)",
       title="Evolution of Global GDP since 2002")+
  guides(color ="none")
```
***
<br>

So, pre-primary school is essential for the development of children, for their development as a children, for their whole education and even for their future adult life. Out-of-school rate for children one year younger than official entry age impact all the education system of countries, the adults they will be and consequently the development of the country itself.

<center>
<span style="color: peru;">**Cut the vicious circle! Act for children!**</span>
</center>

<br>

<right>
“**Education is the most powerful weapon which you can use to change the world**.” Nelson Mandela
</right>
